Evidence for a Third Color Parameter Within the Type~Ia Supernovae of the Nearby Supernova Factory

### 2018-06-16 submit/response.txt

Response 1:

For quick reference we have added a table containing a list of the supernovae and their measurements used in this analysis and a second table listing the supernovae and values of their output parameters.  For completeness, the input data and output Monte Carlo chains are made available in a tar archive on a persistent repository (TODO).  The publicly available spectrophotometric series can be used to check how our pipeline transformation the series into the measures used in this article.

It is unfortunate that the spectrophotometric series of all the supernovae used in this analysis are not yet public.  Nevertheless, there is precedent in astrophysics for interesting results that could stimulate public interest and activity (e.g. verification using independent data sets) to be published before all antecedent data products are made public. This article fits that category.

Despite the time delay pointing to the contrary, there is a concentrated effort within the SNfactory collaboration on making a public data release.  https://drive.google.com/drive/folders/1d52nxfF7eS3HblfbmT3fxgiZod_Z-Gjg?usp=sharing.  DOI number.

@@

Comment 3:

1. It would be useful to show the covariance matrices for all models and discuss how necessary each measurement is. For instance, the Si EW and x1 are highly correlated, and perhaps only one measurement is required.

Response 3:

Note that the correlation/covariance between a subset of parameters are already in the article.  For example the correlation matrix of the per-supernova parameters is given and there is a figure of their distribution.  However, there are too many parameters to make providing their covariance matrix within the article unwieldy.
We did insert out the 35x35 covariance matrix of the global parameters, but it did not have a clean appearance in the manuscript. Instead we provide the Monte Carlo chains in the tar archive from which the covariance of any possible parameter pair can be determined.  The chains are available at  https://drive.google.com/drive/folders/1d52nxfF7eS3HblfbmT3fxgiZod Z-Gjg?usp=sharing.  The

### 2018-06-16 submit/fix.tex
continuing edits

@@

\added{
SNfactory observed supernovae in the redshift range $z \lesssim 0.09$, with a preference for $0.03 < z < 0.08$,
whose first SNIFS observation was projected to be $\lesssim -5$ days before $B$-band peak brightness.
Telescope availability and observing conditions dictated which supernovae eventually would accumulate multiple observations 
over the course of their light curves that would give constraining light curve (e.g.\ SALT2) fits.  At an occurence of $< 10\%$ individual spectral extractions 
}

Our analysis sample is comprised of the
172
supernovae that have been fully processed and have the data coverage to 
give photometric and spectroscopic statistics described above.
The 
spectroscopic-feature measurements, SALT2 light-curve shape parameters  $x_1$,
and their uncertainties
are presented
in \replaced{\citet{2017Chotard}}{Table~\ref{data:tab}. 
For the subset supernovae for whom it is available, we use the host-galaxy mass given in \citep{2013ApJ...770..108C}.

@@

\subsection{Model}
We assume 
that  peak
underlying
 ${\hat{U}}{\hat{B}}\hat{V}\hat{R}\hat{I}$ magnitudes of a supernova  with effects from  astrophysical foregrounds are linearly dependent
on its
 equivalent widths of the Ca~II H\&K and Si~II~$\lambda$4141 spectral features
$EW_{\mathrm{Ca}}$ and $EW_{\mathrm{Si}}$,
and the wavelength of the minimum of 
the Si~II~$\lambda6355$ feature $\lambda_{\mathrm{Si}}$
around $B$-band peak brightness:
these spectral features are associated with SN~Ia  spectroscopic diversity  
\citep{2005ApJ...623.1011B, 2006PASP..118..560B, 2008A&A...492..535A, 2009A&A...500L..17B, 2009PASP..121..238B, 2009ApJ...699L.139W, 2011ApJ...729...55F,2012AJ....143..126B}.
The explicit omission of light-curve shape in our model is compensated by its proxy,
$EW_{\mathrm{Si}}$, at peak brightness
\citep{2008A&A...492..535A, 2011A&A...529L...4C}. 
The underlying magnitudes are also linearly dependent on the per-supernova latent parameters $g_0$ and $g_1$.
Unlike
the spectral parameters ($EW_{\mathrm{Ca}}$, $EW_{\mathrm{Si}}$ and $\lambda_{\mathrm{Si}}$),  $g_0$ and $g_1$ are not directly associated
with observables but rather are inferred as part of the analysis.
A grey magnitude offset, $\Delta$, is included for each supernova
to capture 
potential
band-independent intrinsic dispersion, while also absorbing peculiar-velocity errors introduced when converting
fluxes to luminosities.
With this grey offset the model standardizes colors, not absolute magnitude.
The underlying magnitudes are
\begin{equation}
\begin{pmatrix}
{\hat{U}}\\{\hat{B}}\\{\hat{V}}\\{\hat{R}}\\{\hat{I}}
\end{pmatrix}
=
\Delta \vec{1} +
\begin{pmatrix}
c_{\hat{U}}+\alpha_{\hat{U}} EW_{\mathrm{Ca}} + \beta_{\hat{U}} EW_{\mathrm{Si}} + \eta_{\hat{U}} \lambda_{\mathrm{Si}}  +\gamma^0_{{\hat{U}}} g_0 +\gamma^1_{{\hat{U}}} g_1 \\
c_{\hat{B}}+\alpha_{\hat{B}} EW_{\mathrm{Ca}} + \beta_{\hat{B}} EW_{\mathrm{Si}} + \eta_{\hat{B}} \lambda_{\mathrm{Si}}  +\gamma^0_{{\hat{B}}} g_0 +\gamma^1_{{\hat{B}}} g_1 \\
c_{\hat{V}}+\alpha_{\hat{V}} EW_{\mathrm{Ca}} + \beta_{\hat{V}} EW_{\mathrm{Si}} + \eta_{\hat{V}} \lambda_{\mathrm{Si}} +\gamma^0_{{\hat{V}}} g_0 +\gamma^1_{{\hat{V}}} g_1 \\
c_{\hat{R}}+\alpha_{\hat{R}} EW_{\mathrm{Ca}} + \beta_{\hat{R}} EW_{\mathrm{Si}} + \eta_{\hat{R}} \lambda_{\mathrm{Si}}  +\gamma^0_{{\hat{R}}} g_0 +\gamma^1_{{\hat{R}}} g_1 \\
c_{\hat{I}}+\alpha_{\hat{I}} EW_{\mathrm{Ca}} + \beta_{\hat{I}} EW_{\mathrm{Si}}+ \eta_{\hat{I}} \lambda_{\mathrm{Si}}  +\gamma^0_{{\hat{I}}} g_0 +\gamma^1_{{\hat{I}}} g_1
\end{pmatrix}.
\label{ewsiv:eqn}
\end{equation}

$\vec{1}$ is the Identity vector.
The global parameter vectors\footnote{Global parameters act on each of the five bands: a single component is written with a subscript for the corresponding band, e.g.\  $c_{\hat{U}}$; 
the combination of all components are written as a vector, e.g.\ $\vec{c}=(c_{\hat{U}}, c_{\hat{B}}, c_{\hat{V}}, c_{\hat{R}}, c_{\hat{I}}) $.  The
band subscript or arrow
notationally identify global parameters, in contrast with the per-supernova parameters.
} that describe the SN~Ia population
$\vec{c}$  (units of mag);
$\vec{\alpha}$, $\vec{\beta}$,
$\vec{\eta}$ (units of mag~\AA$^{-1}$), $\vec{\gamma}^0$ and $\vec{\gamma}^1$ (units of mag)
 are the intercept
and slopes
of the linear relationships that
relate
per-supernova
parameters with 
underlying magnitudes.
The notation used in this article is not to be associated with those from previous works.

@@

\subsection{Results for the Latent Parameter Cofactors  $\vec{\gamma}^0$ and $\vec{\gamma}^1$  }
\label{results1:sec}
In this subsection we concentrate on the results for $\vec{\gamma}^0$ and $\vec{\gamma}^1$.
In our model, $g_0$ and $g_1$ are introduced as latent parameters that are related to linear shifts in band magnitudes through
the  coefficients $\vec{\gamma}^0$ and
$\vec{\gamma}^1$.  Otherwise, no physically-motivated suppositions on the properties of $g$ and $\vec{\gamma}$ are imposed.
Since dust extinction is a dominant determinant of supernova colors that is often described by a 2-parameter linear model, we expect
for our two  color terms to capture the effects of dust.
In this section we reexpress our model $\vec{\gamma}$'s in terms of
the dust-extinction model of \citet{1999PASP..111...63F} \citepalias[henceforth referred to as][]{1999PASP..111...63F}.

@@

\subsection{Results for the Latent Parameter Cofactor $\sigma_p \vec{\phi}$}
\label{results2:sec}
The credible intervals of all Model~II global parameters are given in  \S\ref{m12results:sec}.
In this subsection we focus on the effect of adding the third color term  $\sigma_p \vec{\phi}$.
There is strong evidence for a third color parameter.
The 68\% credible interval for $\sigma_p$ is $0.051^{+0.005}_{-0.004}$~mag, highly incompatible with zero. 
The posterior for $\sigma_p \vec{\phi}$ is shown in Figure~\ref{M1ev:fig}.  
Two maxima are allowed due to the sign degeneracy for $\vec{\phi}$.  Practically each of the MCMC chains converges to one of the solutions and does
not migrate between them by crossing zero.  To aid visibility of this figure we thus impose $\phi_I > 0$.  
 The new parameter does affect colors in a  manner distinct from $g_0$ and $g_1$.
A quadratic fraction of
$0.48^{+0.25}_{-0.17}$ of 
 $\vec{\phi}$ lies perpendicular to the $\vec{\gamma}_0$--$\vec{\gamma}_1$ plane.

@@

\subsection{Results}
\subsubsection{Analysis Method and Validation}

@@

\added{
The correlation matrix in Eq.~\ref{corr:eqn} is not close to being singular, meaning that the each of the features conveys
independent supernova variability.  Nevertheless, a strong correlation between two features means that the determination
of one would reduce the residual dispersion in the other.  This information is conveyed in the matrix
\begin{equation}
M =
\begin{pmatrix}
\tiny
\begin{array}{rrrrrrr}
2.76  \pm 0.06 \times 10^{ 1} & 2.75  \pm 0.06 \times 10^{ 1} & 2.67  \pm 0.06 \times 10^{ 1} & 2.75  \pm 0.05 \times 10^{ 1} & 2.74  \pm 0.06 \times 10^{ 1} & 2.71  \pm 0.06 \times 10^{ 1} & 2.75  \pm 0.06 \times 10^{ 1} \\
7.04  \pm 0.16 \times 10^{ 0} & 7.08  \pm 0.16 \times 10^{ 0} & 7.02  \pm 0.16 \times 10^{ 0} & 3.86  \pm 0.19 \times 10^{ 0} & 7.01  \pm 0.16 \times 10^{ 0} & 6.79  \pm 0.21 \times 10^{ 0} & 7.06  \pm 0.16 \times 10^{ 0} \\
1.98  \pm 0.04 \times 10^{ 1} & 2.03  \pm 0.05 \times 10^{ 1} & 2.05  \pm 0.05 \times 10^{ 1} & 2.03  \pm 0.05 \times 10^{ 1} & 2.04  \pm 0.05 \times 10^{ 1} & 2.04  \pm 0.05 \times 10^{ 1} & 2.04  \pm 0.05 \times 10^{ 1} \\
9.92  \pm 0.18 \times 10^{-1} & 5.43  \pm 0.23 \times 10^{-1} & 9.86  \pm 0.19 \times 10^{-1} & 9.95  \pm 0.18 \times 10^{-1} & 9.86  \pm 0.19 \times 10^{-1} & 9.71  \pm 0.26 \times 10^{-1} & 9.92  \pm 0.19 \times 10^{-1} \\
7.33  \pm 0.31 \times 10^{-2} & 7.32  \pm 0.32 \times 10^{-2} & 7.37  \pm 0.32 \times 10^{-2} & 7.32  \pm 0.32 \times 10^{-2} & 7.38  \pm 0.32 \times 10^{-2} & 7.36  \pm 0.33 \times 10^{-2} & 7.36  \pm 0.32 \times 10^{-2} \\
3.82  \pm 0.42 \times 10^{-2} & 3.73  \pm 0.41 \times 10^{-2} & 3.88  \pm 0.43 \times 10^{-2} & 3.79  \pm 0.42 \times 10^{-2} & 3.88  \pm 0.44 \times 10^{-2} & 3.89  \pm 0.44 \times 10^{-2} & 3.88  \pm 0.44 \times 10^{-2} \\
4.95  \pm 1.09 \times 10^{-2} & 4.95  \pm 1.09 \times 10^{-2} & 4.95  \pm 1.09 \times 10^{-2} & 4.95  \pm 1.09 \times 10^{-2} & 4.95  \pm 1.09 \times 10^{-2} & 4.95  \pm 1.09 \times 10^{-2} & 4.97  \pm 1.10 \times 10^{-2} 
\\\end{array}
\end{pmatrix}.
\label{ellipse:eqn}
\end{equation}
For indeces $i, j \in \{EW_{\mathrm{Ca}}, EW_{\mathrm{Si}}, \lambda_{\mathrm{Si}}, x_1, E_{\gamma^0}({\hat{B}}-{\hat{V}}), E_{\gamma^1}({\hat{B}}-{\hat{V}}), A_{p,V}\}$, $M_{ii}$ is the standard deviation in the distribution of parameter $i$, and for $i \ne j$ $M_{ij}$ is the standard deviation in parameter $i$
if parameter $j$ were known perfectly.  For example, for $i= EW_{\mathrm{Si}}$ the standard deviation is 7.08 but when $j=x_1$
is fixed its
standard deviation drops to 3.86.  Therefore, an analysis that considers $x_1$ but not $EW_{\mathrm{Si}}$ would not account for
$3.86\vec{\beta}=\{0.14 ,  0.09,  0.08,  0.07,  0.04 \}$ ${\hat{U}}{\hat{B}}{\hat{V}}{\hat{R}}{\hat{I}}$ magnitude dispersions
that are not intrinsically random.
}

@@

The color-standardization leaves magnitude residuals $\Delta$.
The histogram of the per-supernova  medians of these relative  grey offsets $\Delta-\Delta|_0$,
is shown in Figure~\ref{hist:fig}.  The distribution
has a total standard deviation of
%-----
$0.13$
%-----
mag, and a tail in the positive (fainter) direction.
Supernova peculiar velocities contribute to the standard deviation; of the four
supernovae with the highest values of $\Delta$, two have 
heliocentric
redshifts of 0.0015 and 0.0086. 
\replaced{
Eq.~\ref{corr:eqn} shows $>2 \sigma$ correlations between $\Delta$ and each of $\lambda_{\mathrm{Si}}$, $x_1$, $E_{\gamma^1}({\hat{B}}-{\hat{V}})$, and
$\sigma_p\phi_{\hat{V}}p$, showing that in addition to standardizing colors, these features can also be used to standardize
absolute magnitudes.  }{Recall that model degeneracy between $\Delta$ and the cofactors such as $\alpha$, $\beta$ is removed by enforcing
zero correlation between $\Delta$ and the per-supernova parameters such as $EW_{\mathrm{Ca}}$,  $EW_{\mathrm{Si}}$.
An alternative method to break the degeneracy is to minimize the dispersion in $\Delta$.
}
The focus of this article is color standardization so we defer consideration of absolute-magnitude standardization to future work.

>>>>

\added{The current SNfactory sample consists of 198 SNe~Ia
having fully-processed spectrophotometric lightcurve data, including
observations on at least two photometric nights, final references, and a
host spectroscopic redshift. 
These supernovae have well-sampled lightcurves as defined by having
at minimum 5 spectra while still active, with at least four within
epochs  $[-10, 35]$ days, at least one within $[-10, 7]$ days, at least one within $[7, 20]$ days,
and at least one color (Anyone know what this means? See slide 3 of
\url{https://snf-doc.lbl.gov/twiki/pub/Conf/VideoConf20160317/IDR_03-17-2016.pdf}) within $[-8, 10]$ days of peak $B$ brightness.
Synthetic-photometric lightcurves derived from the spectra are fit to the SALT2~(v2.4)
model; we require that supernovae have no more than 20\% points that are
$>0.2$ mag from the best-fit model, and require a Normalized Median Absolute Deviation (nMAD) of residuals
of  $< 0.12$ mag.
%These all have at least 5 spectra while the SN is
%active, and pass the quality cuts suggested by \citet{2010A&A...523A...7G}.
More than 80\% of our SNe are from searches where there was no
pre-selection based on host galaxy properties (those
whose names start with ``SNF'', ``LSQ'', or ``PTF'' in Table~\ref{data:tab}).

%Quality criteria for a 'good' lightcurve are the following: At least 5 nights of observation
%No more than 20\% of rejection (after outlier cut>0.2mag) nMAD of residuals < 0.12 mag
%Phase coverage (4 epochs in [-10 < p < +35] days, 1 epochs in [-10 < p < +7] days, 1 epochs in [+7 < p < +20] days,1 color in [-8 < p < +10] days)

From that sample, we further require at least one spectrum with $\pm 2.5$ days of B-band peak brightness.
We exclude six SNe~Ia in the
SN~1991T, SN~1991bg and SN~2002cx subclasses,
SNF20080905-005 described in \citet{leget:thesis} as
being spectroscopically peculiar, and
the extremely red SN2012cu  who has a SALT2 $C \approx 1$ \citep{2017ApJ...836..157H} that is several times larger than the colors
of in-sample supernovae.

For the purposes of this article, it is not important whether our sample population is representative
of the underlying population.  This article seeks
underlying parameters that
affect SN~Ia colors; any such parameters found in a subsample are still
relevant parameters when dealing with the population as a whole.

There is a fraction of SNfactory-observed supernovae that do have the
observations to satisfy the well-sampled light-curve criteria, but are excluded due to some observations
having not completed full processing.  Incomplete processing can occur for data taken during poor
observing conditions, when there are instrumental artifacts, and late in a supernova's
light-curve when the signal-to-noise and contrast relative to the host are low.  The sample-selection exclusion of these supernovae are not related,
at least directly, to the at-maximum observables used in this analysis.  Our results are insensitive to potential sample-selection
bias due to stochasticity in the measurements.
}

@@

Our analysis sample is comprised of the
172
supernovae that have been fully processed and have the data coverage to 
give photometric and spectroscopic statistics described above.
The 
spectroscopic-feature measurements, SALT2 light-curve shape parameters  $x_1$,
and their uncertainties
are presented
in \replaced{\citet{2017Chotard}}{Table~\ref{data:tab}. 
For the subset supernovae for whom it is available, we use the host-galaxy mass given in \citet{2013ApJ...770..108C}.

@@

\added{It is important to note that the model and those in subsequent sections are applied to our supernova sample;
the ensuing results thus describe our SN~Ia sample's underlying population. }

\subsection{Model}
We assume 
that  peak
underlying
 ${\hat{U}}{\hat{B}}\hat{V}\hat{R}\hat{I}$ magnitudes of a supernova  with effects from  astrophysical foregrounds are linearly dependent
on its
 equivalent widths of the Ca~II H\&K and Si~II~$\lambda$4141 spectral features
$EW_{\mathrm{Ca}}$ and $EW_{\mathrm{Si}}$,
and the wavelength of the minimum of 
the Si~II~$\lambda6355$ feature $\lambda_{\mathrm{Si}}$
around $B$-band peak brightness:
these spectral features are associated with SN~Ia  spectroscopic diversity  
\citep{2005ApJ...623.1011B, 2006PASP..118..560B, 2008A&A...492..535A, 2009A&A...500L..17B, 2009PASP..121..238B, 2009ApJ...699L.139W, 2011ApJ...729...55F,2012AJ....143..126B}.
The explicit omission of light-curve shape in our model is compensated by its proxy,
$EW_{\mathrm{Si}}$, at peak brightness
\citep{2008A&A...492..535A, 2011A&A...529L...4C}. 
The underlying magnitudes are also linearly dependent on the per-supernova latent parameters $g_0$ and $g_1$.
Unlike
the spectral parameters ($EW_{\mathrm{Ca}}$, $EW_{\mathrm{Si}}$ and $\lambda_{\mathrm{Si}}$),  $g_0$ and $g_1$ are not directly associated
with observables but rather are inferred as part of the analysis.
A grey magnitude offset, $\Delta$, is included for each supernova
to capture 
potential
band-independent intrinsic dispersion, while also absorbing peculiar-velocity errors introduced when converting
fluxes to luminosities.
With this grey offset the model standardizes colors, not absolute magnitude.
The underlying magnitudes are
\begin{equation}
\begin{pmatrix}
{\hat{U}}\\{\hat{B}}\\{\hat{V}}\\{\hat{R}}\\{\hat{I}}
\end{pmatrix}
=
\Delta \vec{1} +
\begin{pmatrix}
c_{\hat{U}}+\alpha_{\hat{U}} EW_{\mathrm{Ca}} + \beta_{\hat{U}} EW_{\mathrm{Si}} + \eta_{\hat{U}} \lambda_{\mathrm{Si}}  +\gamma^0_{{\hat{U}}} g_0 +\gamma^1_{{\hat{U}}} g_1 \\
c_{\hat{B}}+\alpha_{\hat{B}} EW_{\mathrm{Ca}} + \beta_{\hat{B}} EW_{\mathrm{Si}} + \eta_{\hat{B}} \lambda_{\mathrm{Si}}  +\gamma^0_{{\hat{B}}} g_0 +\gamma^1_{{\hat{B}}} g_1 \\
c_{\hat{V}}+\alpha_{\hat{V}} EW_{\mathrm{Ca}} + \beta_{\hat{V}} EW_{\mathrm{Si}} + \eta_{\hat{V}} \lambda_{\mathrm{Si}} +\gamma^0_{{\hat{V}}} g_0 +\gamma^1_{{\hat{V}}} g_1 \\
c_{\hat{R}}+\alpha_{\hat{R}} EW_{\mathrm{Ca}} + \beta_{\hat{R}} EW_{\mathrm{Si}} + \eta_{\hat{R}} \lambda_{\mathrm{Si}}  +\gamma^0_{{\hat{R}}} g_0 +\gamma^1_{{\hat{R}}} g_1 \\
c_{\hat{I}}+\alpha_{\hat{I}} EW_{\mathrm{Ca}} + \beta_{\hat{I}} EW_{\mathrm{Si}}+ \eta_{\hat{I}} \lambda_{\mathrm{Si}}  +\gamma^0_{{\hat{I}}} g_0 +\gamma^1_{{\hat{I}}} g_1
\end{pmatrix}.
\label{ewsiv:eqn}
\end{equation}
$\vec{1}$ is the Identity vector.
The global parameter vectors\footnote{Global parameters act on each of the five bands: a single component is written with a subscript for the corresponding band, e.g.\  $c_{\hat{U}}$; 
the combination of all components are written as a vector, e.g.\ $\vec{c}=(c_{\hat{U}}, c_{\hat{B}}, c_{\hat{V}}, c_{\hat{R}}, c_{\hat{I}}) $.  The
band subscript or arrow
notationally identify global parameters, in contrast with the per-supernova parameters.
} that describe the SN~Ia population
$\vec{c}$  (units of mag);
$\vec{\alpha}$, $\vec{\beta}$,
$\vec{\eta}$ (units of mag~\AA$^{-1}$), $\vec{\gamma}^0$ and $\vec{\gamma}^1$ (units of mag)
 are the intercept
and slopes
of the linear relationships that
relate
per-supernova
parameters with 
underlying magnitudes.
The notation used in this article is not to be associated with those from previous works.

@@

\subsection{Results for the Latent Parameter Cofactors  $\vec{\gamma}^0$ and $\vec{\gamma}^1$  }
\label{results1:sec}
\added{The MCMC chain for all parameters are accessible from a DOI link \citep{colordata}.}
In this subsection we concentrate on the results for $\vec{\gamma}^0$ and $\vec{\gamma}^1$.
In our model, $g_0$ and $g_1$ are introduced as latent parameters that are related to linear shifts in band magnitudes through
the  coefficients $\vec{\gamma}^0$ and
$\vec{\gamma}^1$.  Otherwise, no physically-motivated suppositions on the properties of $g$ and $\vec{\gamma}$ are imposed.
Since dust extinction is a dominant determinant of supernova colors that is often described by a 2-parameter linear model, we expect
for our two  color terms to capture the effects of dust.
In this section we reexpress our model $\vec{\gamma}$'s in terms of
the dust-extinction model of \citet{1999PASP..111...63F} \citepalias[henceforth referred to as][]{1999PASP..111...63F}.

@@

\subsection{Results for the Latent Parameter Cofactor $\sigma_p \vec{\phi}$}
\label{results2:sec}
\added{The MCMC chain for all parameters are accessible from a DOI link \citep{colordata}.}
The credible intervals of all Model~II global parameters are given in  \S\ref{m12results:sec}.
In this subsection we focus on the effect of adding the third color term  $\sigma_p \vec{\phi}$.
There is strong evidence for a third color parameter.
The 68\% credible interval for $\sigma_p$ is $0.051^{+0.005}_{-0.004}$~mag, highly incompatible with zero. 
The posterior for $\sigma_p \vec{\phi}$ is shown in Figure~\ref{M1ev:fig}.  
Two maxima are allowed due to the sign degeneracy for $\vec{\phi}$.  Practically each of the MCMC chains converges to one of the solutions and does
not migrate between them by crossing zero.  To aid visibility of this figure we thus impose $\phi_I > 0$.  
 The new parameter does affect colors in a  manner distinct from $g_0$ and $g_1$.
A quadratic fraction of
$0.48^{+0.25}_{-0.17}$ of 
 $\vec{\phi}$ lies perpendicular to the $\vec{\gamma}_0$--$\vec{\gamma}_1$ plane.

@@

\subsection{Results}
\added{The MCMC chain for all parameters are accessible from a DOI link \citep{colordata}.}

\subsubsection{Analysis Method and Validation}

@@

\added{
The correlation matrix in Eq.~\ref{corr:eqn} is not close to being singular, meaning that the each of the features conveys
independent supernova variability.  Nevertheless, a strong correlation between two features means that the determination
of one would reduce the residual dispersion in the other.  The amount of residual dispersion is conveyed in the matrix
\begin{equation}
M =
\begin{pmatrix}
\tiny
\begin{array}{rrrrrrr}
2.76  \pm 0.06 \times 10^{ 1} & 2.75  \pm 0.06 \times 10^{ 1} & 2.67  \pm 0.06 \times 10^{ 1} & 2.75  \pm 0.05 \times 10^{ 1} & 2.74  \pm 0.06 \times 10^{ 1} & 2.71  \pm 0.06 \times 10^{ 1} & 2.75  \pm 0.06 \times 10^{ 1} \\
7.04  \pm 0.16 \times 10^{ 0} & 7.08  \pm 0.16 \times 10^{ 0} & 7.02  \pm 0.16 \times 10^{ 0} & 3.86  \pm 0.19 \times 10^{ 0} & 7.01  \pm 0.16 \times 10^{ 0} & 6.79  \pm 0.21 \times 10^{ 0} & 7.06  \pm 0.16 \times 10^{ 0} \\
1.98  \pm 0.04 \times 10^{ 1} & 2.03  \pm 0.05 \times 10^{ 1} & 2.05  \pm 0.05 \times 10^{ 1} & 2.03  \pm 0.05 \times 10^{ 1} & 2.04  \pm 0.05 \times 10^{ 1} & 2.04  \pm 0.05 \times 10^{ 1} & 2.04  \pm 0.05 \times 10^{ 1} \\
9.92  \pm 0.18 \times 10^{-1} & 5.43  \pm 0.23 \times 10^{-1} & 9.86  \pm 0.19 \times 10^{-1} & 9.95  \pm 0.18 \times 10^{-1} & 9.86  \pm 0.19 \times 10^{-1} & 9.71  \pm 0.26 \times 10^{-1} & 9.92  \pm 0.19 \times 10^{-1} \\
7.33  \pm 0.31 \times 10^{-2} & 7.32  \pm 0.32 \times 10^{-2} & 7.37  \pm 0.32 \times 10^{-2} & 7.32  \pm 0.32 \times 10^{-2} & 7.38  \pm 0.32 \times 10^{-2} & 7.36  \pm 0.33 \times 10^{-2} & 7.36  \pm 0.32 \times 10^{-2} \\
3.82  \pm 0.42 \times 10^{-2} & 3.73  \pm 0.41 \times 10^{-2} & 3.88  \pm 0.43 \times 10^{-2} & 3.79  \pm 0.42 \times 10^{-2} & 3.88  \pm 0.44 \times 10^{-2} & 3.89  \pm 0.44 \times 10^{-2} & 3.88  \pm 0.44 \times 10^{-2} \\
4.95  \pm 1.09 \times 10^{-2} & 4.95  \pm 1.09 \times 10^{-2} & 4.95  \pm 1.09 \times 10^{-2} & 4.95  \pm 1.09 \times 10^{-2} & 4.95  \pm 1.09 \times 10^{-2} & 4.95  \pm 1.09 \times 10^{-2} & 4.97  \pm 1.10 \times 10^{-2} 
\\\end{array}
\end{pmatrix}.
\label{ellipse:eqn}
\end{equation}
For indeces $i, j \in \{EW_{\mathrm{Ca}}, EW_{\mathrm{Si}}, \lambda_{\mathrm{Si}}, x_1, E_{\gamma^0}({\hat{B}}-{\hat{V}}), E_{\gamma^1}({\hat{B}}-{\hat{V}}), A_{p,V}\}$, $M_{ii}$ is the standard deviation of the parameter $i$ distribution, and for $i \ne j$ $M_{ij}$ is the standard deviation in parameter $i$
if parameter $j$ were known perfectly.  For example, for $i= EW_{\mathrm{Si}}$ the standard deviation is 7.08 but when $j=x_1$
is fixed its
standard deviation drops to 3.86.  Therefore, an analysis that considers $x_1$ but not $EW_{\mathrm{Si}}$ would not account for
$3.86\vec{\beta}=\{0.14 ,  0.09,  0.08,  0.07,  0.04 \}$ ${\hat{U}}{\hat{B}}{\hat{V}}{\hat{R}}{\hat{I}}$ magnitude dispersions
that are not intrinsically random.
}

@@

The color-standardization leaves magnitude residuals $\Delta$.
The histogram of the per-supernova  medians of these relative  grey offsets $\Delta-\Delta|_0$,
is shown in Figure~\ref{hist:fig}.  The distribution
has a total standard deviation of
%-----
$0.13$
%-----
mag, and a tail in the positive (fainter) direction.
Supernova peculiar velocities contribute to the standard deviation; of the four
supernovae with the highest values of $\Delta$, two have 
heliocentric
redshifts of 0.0015 and 0.0086. 
\replaced{
Eq.~\ref{corr:eqn} shows $>2 \sigma$ correlations between $\Delta$ and each of $\lambda_{\mathrm{Si}}$, $x_1$, $E_{\gamma^1}({\hat{B}}-{\hat{V}})$, and
$\sigma_p\phi_{\hat{V}}p$, showing that in addition to standardizing colors, these features can also be used to standardize
absolute magnitudes.  }{Recall that the model degeneracy between $\Delta$ and the cofactors (e.g.\ $\alpha$, $\beta$) is removed by enforcing
zero correlation between $\Delta$ and the per-supernova parameters (e.g.\ $EW_{\mathrm{Ca}}$,  $EW_{\mathrm{Si}}$).
An alternative method to break the degeneracy is to minimize the dispersion in $\Delta$.
}
The focus of this article is color standardization so we defer consideration of absolute-magnitude standardization to future work.
